
\begin{thebibliography}{99}
{\fontfamily{lmtt}\selectfont\scriptsize


\bibitem{RaubalAdams}
Benjamin Adams and Martin Raubal, 
\cq{A Metric Conceptual Space Algebra}.
\biburl{https://pdfs.semanticscholar.org/521a/cbab9658df27acd9f40bba2b9445f75d681c.pdf}

\bibitem{RaubalAdamsCSML}
Benjamin Adams and Martin Raubal, 
\cq{Conceptual Space Markup Language (CSML): Towards the Cognitive Semantic Web}.
\biburl{http://idwebhost-202-147.ethz.ch/Publications/RefConferences/ICSC_2009_AdamsRaubal_Camera-FINAL.pdf}


\bibitem{AsherPustejovsky}
Nicholas Asher and James Pustejovsky, 
\cq{A Type Composition Logic for Generative Lexicon}
\biburl{https://www.cs.brandeis.edu/~jamesp/classes/cs216-2009/readings2009/TCLforGL.pdf}


\bibitem{BarkerShanTG}
Chris Barker and Chung-Chieh Shan, 
\cq{Types as Graphs: Continuations in Type Logical Grammar}
\biburl{https://www.nyu.edu/projects/barker/barker-shan-types-as-graphs.pdf}

\bibitem{BernardyEtAl}
Jean-Philippe Bernardy, et. al.,
\cq{Parametricity and Dependent Types}.
\biburl{http://www.staff.city.ac.uk/~ross/papers/pts.pdf}

\bibitem{BittnerSmithDonnelly}
Thomas Bittner, Barry Smith, and Maureen Donnelly,
\cq{The logic of systems of granular partitions.}
\biburl{http://ontology.buffalo.edu/smith/articles/BittnerSmithDonnelly.pdf}

\bibitem{LineBrandt}
Line Brandt, 
\cq{The Communicative Mind.}
Newcastle-upon-Tyne: Cambridge Scholars Publishing, 2013

%
\bibitem{Belohlavek}
Radim B\v{e}lohl\'avek and Vladim{\i\OldI}r Sklen\'a\v{r}, 
\cq{Formal Concept Analysis Constrained by Attribute-Dependency Formulas}
B. Ganter and R. Godin, eds.,: ICFCA 2005, LNCS 3403, pp. 176-191, Berlin, Springer-Verlag, 2005.
\biburl{http://belohlavek.inf.upol.cz/publications/BeSk_Fcacadf.pdf}

%
\bibitem{PerAageBrandt}
Per Aage Brandt, 
\cq{Spaces, Domains, and Meaning: Essays in Cognitive Semiotics}.
Newcastle-upon-Tyne: Cambridge Scholars Publishing, 2013
\biburl{http://semiotics.au.dk/fileadmin/Semiotics/pdf/per-aage-brandt/Per_Aage_Brandt__Spaces__Domains__and_Meaning._Essays_in_Co.pdf}


\bibitem{ChatzikyriakidisLuo}
Stergios Chatzikyriakidis and Zhaohui Luo, 
\cq{Individuation Criteria, Dot-types and Copredication: A View from Modern Type Theories.}
\intitle{Association for Computational Linguistics, Proceedings of the 14th 
Meeting on the Mathematics of Language (MoL 14)}, pp. 39–50, 2015. 
\biburl{http://www.aclweb.org/anthology/W15-2304}

\bibitem{ChoeCharniak}
Do Kook Choe and Eugene Charniak, 
\cq{Parsing as Language Modeling}.
\tinyurl{https://aclweb.org/anthology/D16-1257}


\bibitem{InteractingConceptualSpaces}
Bob Coecke, \i{et. al.}, 
\cq{Interacting Conceptual Spaces I:
Grammatical Composition of Concepts}.
\biburl{https://arxiv.org/pdf/1703.08314.pdf}


\bibitem{Zenker}
Peter \Gardenfors{} and Frank Zenker,  
\cq{Theory Change as Dimensional Change: Conceptual Spaces 
	Applied to the Dynamics of Empirical Theories}.
\intitle{Synthese 190(6)}, pp. 1039-1058, 2013.  
\biburl{http://lup.lub.lu.se/record/1775234}

\bibitem{KennethHolmqvist}
Kenneth Holmqvist,
\cq{Conceptual Engineering: Implementing cognitive semantics}, 
in Jens Allwood and Peter \Gardenfors, eds., 
\intitle{Cognitive Semantics}, pp 153 - 171, Amsterdam,
Philadelphia: John Benjamins, 1999.

\bibitem{HolmqvistDiss}
Kenneth Holmqvist, 
\cq{Implementing Cognitive Semantics: Image schemata, 
valence accommodation, and valence suggestion for  
AI and computational linguistics}.
PhD thesis, Dept. of Cognitive Science, Lund University, Lund, Sweden, 1993.


\bibitem{Johnson}
Mark Johnson, \cq{The Body in the Mind: 
The Bodily Basis of Meaning, Imagination, and Reason}.  %Chicago, 1990
University of Chicago Press, 1990


\bibitem{KubotaLevine}
Yusuke Kubota and Robert Levine,
\cq{The syntax-semantics interface of \sq{respective} predication:
A unified analysis in Hybrid Type-Logical Categorial Grammar}
\biburl{https://www.asc.ohio-state.edu/levine.1/publications/kl-resp.pdf}


\bibitem{LakoffJohnson}
George Lakoff and Mark Johnson, 
\cq{Philosophy in the Flesh: the Embodied Mind and its Challenge to Western Thought}.
New York: Basic Books, 1999.

\bibitem{Langacker87}
Ronald Langacker, 
\cq{Nouns and Verbs}
\intitle{Language}, Vol. 63, No. 1 (March 1987) pp. 53-94.

\bibitem{LangackerFoundations}
Ronald Langacker, 
\cq{Foundations of Cognitive Grammar, vol. 1}.  
Stanford University Press, 1991

\bibitem{ZhaohuiLuo}
Zhaohui Luo,
\cq{Type-Theoretical Semantics with Coercive Subtyping}.
\biburl{https://www.cs.rhul.ac.uk/home/zhaohui/ESSLLI11notes.pdf}

\bibitem{LuoSoloviev}
Zhaohui Luo and Sergei Soloviev,
\cq{Dependent Coercions}.
\biburl{https://www.sciencedirect.com/science/article/pii/S1571066105803147}

\bibitem{MeryMootRetore}
Bruno Mery, Richard Moot, and Christian Retor\'e, \cq{
	Plurals: Individuals and sets
	in a richly typed semantics}.  
\biburl{http://arxiv.org/pdf/1401.0660.pdf} \archiveDate{3 Jan 2014}


\bibitem{EstherPascual}
Esther Pascual, 
\cq{Fictive Interaction: The conversation frame in thought, language, and discourse}.
Philadelphia, John Benjamins, 2014.

\bibitem{EfePeker}
Efe Peker,
\cq{Following 9/11: George W. Bush's Discursive Re-Articulation of American Social Identity}
Master's Dissertation, Link\"oping University, Sweden
\biburl{http://www.diva-portal.org/smash/get/diva2:21407/FULLTEXT01.pdf}

\bibitem{PetitotSyntaxe}
Jean Petitot, \cq{Syntax Topologique et Grammaire Cognitive}.
\intitle{Langages 25.103}, pp. 97-128, 1991. 

%?
\bibitem{JeanPetitot}
Jean Petitot, \cq{The morphodynamical turn of cognitive linguistics}.
\biburl{https://journals.openedition.org/signata/549}

\bibitem{SivaReddy}
Siva Reddy, \i{et. al},
\cq{Transforming Dependency Structures to Logical Forms for Semantic Parsing}
\biburl{https://aclweb.org/anthology/Q16-1010}

\bibitem{TerryRegier}
Terry Regier, 
\i{The Human Semantic Potential: 
Spatial Language and Constrained Connectionism}.
Cambridge, MA: MIT Press, 2014.

\bibitem{MattSelway}
Matt Selway, \i{et. al},
\cq{Configuring Domain Knowledge for Natural Language Understanding}
\biburl{http://ceur-ws.org/Vol-1128/paper9.pdf}


\bibitem{KiyoshiSudo}
Kiyoshi Sudo, \i{et. al},
\cq{An Improved Extraction Pattern Representation Model
for Automatic IE Pattern Acquisition}
\biburl{https://www.aclweb.org/anthology/P03-1029}


\bibitem{TanakaEtAl}
Ribeka Tanaka, et. al., \cq{Factivity and Presupposition
in Dependent Type Semantics}
\biburl{http://www.lirmm.fr/tytles/Articles/Tanaka.pdf}

\bibitem{OrlinVakarelov}
Orlin Vakarelov,
\cq{Pre-cognitive Semantic Information}.
\biburl{https://link.springer.com/article/10.1007/s12130-010-9109-5}

\bibitem{VakarelovAgent}
Orlin Vakarelov, 
\cq{The cognitive agent: Overcoming
informational limits}
\biburl{https://philarchive.org/archive/VAKTCAv1}

%\bibitem{WiegandMereology}
\bibitem{OlavKWiegand}
Olav K. Wiegand, 
\cq{A Formalism Supplementing Cognitive Semantics Based on a New Approach to Mereology}.
Ingvar Johansson, Bertin Klein and Thomas Roth-Berghofer, eds.,  
\intitle{Contributions to the Third International Workshop on Philosophy and Informatics}, 2006. 

\bibitem{WiegandGestalts}
\underlines
\cq{On referring to Gestalts}.
Mirja Hartimo, ed., \intitle{Phenomenology and Mathematics}, pp. 183-211.  Springer, 2010. 

\bibitem{Wille}
Rudolf Wille, 
\cq{Conceptual Graphs and Formal Concept Analysis}.
\intitle{Conceptual Structures: Fulfilling Peirce's Dream
Fifth International Conference on Conceptual Structures}, 1997.

\bibitem{JordanZlatev}
Jordan Zlatev, 
%\underlines
\cq{The Dependence of Language on Consciousness}.
\biburl{http://www.mrtc.mdh.se/~gdc01/work/ARTICLES/2014/4-IACAP\%202014/IACAP14-GDC/pdf/Language-Consciousness-Zlatev.pdf}


}
\end{thebibliography}
