
\begin{frame}{\ft{NA3 Business Strategy}}
\section{Business Strategy}
\vspace{2pt}	

{\begin{minipage}[c]{1.01\textwidth}	
{\color[rgb]{.0,0.1,0.1}{\Large\fontfamily{bch}\fontseries{b}\selectfont The NA3 Protocols encapsulate new 
technical models for Qt development 
and address recognized gaps in the 
Qt ecosystem, particularly the 
lack of a standard Qt Cloud Services.  
The strategy for promoting NA3 is therefore 
oriented to establishing NA3 within the 
Qt community, using the NA3's Qt implementation 
as a foundation to 
document NA3's essential hypergraph and type-theoretic 
concepts.  On this basis NA3 can then be promoted in 
other application-development contexts.}}
%\vspace{1em}	
%Combining Natural Language Processing and Conversation Analysis: \\
%for Next Generation Language and Conversation Tools
\end{minipage}}
%\vspace{1em}
	
\vspace{15pt}	
	
\definecolor{blueback}{RGB}{0,100,100}	
\definecolor{bluefront}{RGB}{0,100,50}
	
{\Large\fontfamily{uhv}\selectfont
\begin{minipage}{\textwidth}
\fcolorbox{darkRed}{blueback}{
\begin{minipage}{.98\textwidth}
\vspace*{1pt}	
\hspace{-2pt}%
\fcolorbox{darkRed}{white}%
{\begin{minipage}{.99\textwidth}%
\vspace{.8em}
{\centerline{\LARGE\color{bluefront} \textbf{Within the Qt Market}}}
\vspace{1em}
{\setlength{\leftmargini}{30pt}
{\Large\begin{itemize}
\sqitem {\lsep} Promote NCN as a 
standard solution for Qt/Cloud Integration.
\vspace{.2em}
\sqitem {\lsep} Promote A3R tools 
for building custom scripting languages for Qt.
\vspace{.2em}
\sqitem {\lsep} \parbox[t]{18.8cm}{Promote the A3R protocol 
as a standard model for inter-application networking, 
describing applications, and 
serializing application-specific data structures.}
\vspace{.2em}
\sqitem {\lsep} \parbox[t]{20.5cm}{On the basis 
of these enhancements to the Qt ecosystem, LTS hopes to join the Qt partners program, 
which would expose NA3's unique features to a 
worldwide developer community.}

\end{itemize}}}\vspace{3pt}
\end{minipage}}\\\vspace{10pt}\\
%
%\color{white}{\hrule}
\vspace*{-41pt}
\hspace{-6pt}\fcolorbox{darkRed}{white}{\begin{minipage}{.99\textwidth}%
\vspace{.8em}
{\centerline{\LARGE\color{bluefront}  \textbf{Outside of Qt (see slide 9)}}\vspace{.8em}}
{\setlength{\leftmargini}{30pt}
{\Large%\begin{minipage}{.96\textwidth}
\begin{itemize}
\sqitem {\lsep} Generalize the NA3 C++ reflection model and 
hypergraph libraries to standard (non-Qt) C++.
\vspace{.5em}  
\sqitem {\lsep} Implement the A3R Protocols for standard 
C++ and for other languages (C\#, Java, etc.).
\vspace{.5em}  
\sqitem {\lsep} \parbox[t]{20cm}{Implement language-agnostic 
hypergraph serialization to allow 
A3R networking between applications written
for different operating systems and/or programming languages.}
 \vspace{.8em}
\end{itemize}%\end{minipage}
}
}
\end{minipage}}\vspace{3em}\end{minipage}}
\end{minipage}}
\end{frame}
